Quality measure notes

### 2021-08-27 StateOfPractice/Ao-Report/15InterviewsWithDevelopers.tex
MI report - updated ch 6 and more
MI report - updated ch 5

We applied our method to the MI domain (Section \ref{sec_applying_method}) and summarize some answers from the interviews with developers in this section. We only discuss the answers that we think are the most exciting and essential in this section, and summarize the rest in Appendix \ref{ap_interview}.

We interviewed nine developers from eight of the 29 MI software projects. The eight projects are \textit{3D Slicer}, \textit{INVESALIUS 3}, \textit{dwv}, \textit{BioImage Suite Web}, \textit{ITK-SNAP}, \textit{MRIcroGL}, \textit{Weasis}, and \textit{OHIF}. We spent about 90 minutes for each interview and asked 20 prepared questions. We also asked following-up questions when we thought it was worth diving deeper. One participant was too busy to have an interview, so they wrote down their answers. The interviewees may have provided multiple answers to each question. Thus, when counting the number of answers, the total result can be larger than nine.

@@

By asking questions 9, 10, and 12, we tried to identify the pain points during the development process in the eight projects. The pain points include current and past obstacles. We also asked the interviewees how they would solve the problems. Questions 9, 10, and 12 are as follows:

\begin{description}
\item[Q9.] Currently, what are the most significant obstacles in your development process?
\item[Q10.] How might you change your development process to remove or reduce these obstacles?
\item[Q12.] In the past, is there any major obstacle to your development process that has been solved? How did you solve it?
\end{description}

@@

\begin{table}[H]
\centering
\begin{tabular}{llll}
\hline
\multirow{2}{*}{Grp} & \multirow{2}{*}{Obstacle} & \multicolumn{2}{l}{Num ans.} \\ \cline{3-4} 
 &  & current & past \\ \hline
\multirow{2}{*}{1} & Lack of fundings & 3 &  \\
 & Lack of time to devote to the project & 2 & 1 \\ \hdashline
\multirow{3}{*}{2} & Hard to keep up with changes in OS and libraries & 1 &  \\
 & Hard to support multiple OS & 2 &  \\
 & Hard to support lower-end computers & 1 & 2 \\ \hdashline
3 & Lack of access to real-world datasets for testing & 3 & 2 \\ \hdashline
\multirow{7}{*}{Oths}
 & Hard to have a high level roadmap from the start & 1 &  \\
 & Not enough participants for usability tests & 1 &  \\
 & Only a few people fully understand the large codebase & 1 &  \\
 & Hard to transfer to new technologies & & 2 \\
 & Hard to understand users' needs & & 1 \\
 & Hard to maintain good documentations & & 1 \\ \hline
\end{tabular}
\caption{\label{tab_obstacles}Current and past obstacles by the numbers of interviewees with the answers}
\end{table}

The interviewees provided some potential and proven solutions for the problems in Table \ref{tab_obstacles}. We group these pain points into three major groups, and put the less mentioned ones into the group \textit{Others}. Sections \ref{sec_pain_points_1}, \ref{sec_pain_points_2}, and \ref{sec_pain_points_3} include further discussion about the three major groups of pain points and their solutions.

\subsection{Pain Points in Group 1}
\label{sec_pain_points_1}
We summarize the pain points in Group 1 as
\textbf{the lack of fundings and time.}

We also summarize the potential and proven solutions as follows.

@@

\subsection{Pain Points in Group 2}
\label{sec_pain_points_2}
We summarize the pain points in Group 2 as
\textbf{the difficulty to balance between four factors: cross-platform compatibility, convenience to development \& maintenance, performance, and security.} They are also related to the choice between native application and web application.

We also summarize the potential and proven solutions as follows.

@@

The interviewees did not conclude any perfect solutions.

\subsection{Pain Points in Group 3}
\label{sec_pain_points_3}
The pain point in Group 3 is
\textbf{the lack of access to real-world datasets for testing.}

We also summarize the potential and proven solutions as follows.

@@

\noindent\textit{Proven solutions from interviewees:}
\begin{itemize}
\item asking the users to provide deidentified copies of medical images if they have problems loading the images;
\item sending the beta versions of software to medical workers who can access the data and complete the tests;
\item if (part of) the team belongs to a medical school or a hospital, using the datasets they can access to;
\item if the team has access to MRI scanners, self-building MI datasets;
\item if the team has connections with MI equipment manufacturers, asking for their help on data format problems;
\item storing all images that cause special problems, and maintaining this special dataset over time.
\end{itemize}

@@

Table \ref{tab_opinion_doc} summarizes interviewees' opinions on documentation. Interviewees from each of the eight projects thought that documentation was important to their projects, and most of them said that it could save their time to answer questions from users and developers. However, most of them saw the need to improve their documentation, and only three of them thought that their documentations conveyed information clearly enough.

\begin{table}[H]
\centering
\begin{tabular}{ll}
\hline
Opinion on documentation & Num ans. \\ \hline
Documentation is vital to the project& 8 \\
Documentation of the project needs improvements & 7 \\
Referring to documentation saves time to answer questions & 6 \\
Lack of time to maintain good documentation & 4 \\
Documentation of the project conveys information clearly & 3 \\
Coding is more preferable than documentation & 2 \\
Users help each other by referring to documentation & 1 \\ \hline
\end{tabular}
\caption{\label{tab_opinion_doc}Opinions on documentation by the numbers of interviewees with the answers}
\end{table}

Table \ref{tab_doc_tools} lists some of the tools and methods mentioned by the interviewees, which they used for documentation.

@@

\section{Contribution Management and Project Management}
\label{sec_contribution_pm}
We tried to understand how the teams managed the contributions and their projects by asking the questions:

\begin{description}
\item[Q5.] Do you have a defined process for accepting new contributions into your team?
\item[Q13.] What is your software development model? For example, waterfall, agile, etc.
\item[Q14.] What is your project management process? Do you think improving this process can tackle the current problem? Were any project management tools used?
\end{description}

\textit{}

\textit{}

\textit{}

Maybe some team had a documented process for accepting new contributions, but none talked about it during the interview. However, most of them mentioned using GitHub and pull requests to manage contributions from the community. The interviewees generally gave very positive feedback on using GitHub. Some also said they had handled the project repository with some other tools, and eventually transferred to git and GitHub. Table \ref{tab_method_receive_contributions} shows the number of times the interviewees mentioned the methods of receiving contributions.

@@

Additionally, the \textit{3D Slicer} team encouraged users to develop their extensions for specific use cases, and the \textit{OHIF} team was trying to enable the use of plug-ins; the interviewee from the \textit{ITK-SNAP} team said one way of accepting new team members was through funded academic projects.

@@

Most interviewees skipped the question ``Do you think improving this process can tackle the current problem?". We concluded that a yes-or-no question is not very informative here. The interviewee from the \textit{OHIF} team gave a positive answer to this question. They believed that a better project management process can improve the efficiency of junior developers. They also improved the project management tools (from public Jira to public GitHub repository plus private Jira), so they could better communicate externally and internally.

\section{Discussions on Software Qualities}
\label{sec_interview_software_qualities}
Questions 15, 16, 17, 18, and 20 are about software qualities \textit{correctness}, \textit{maintainability}, \textit{understandability}, \textit{usability}, and \textit{reproducibility} respectively. We asked these questions to understand the threats to these qualities and the developers' strategies to improve them.

\subsection{Correctness}

\begin{description}
\item[Q15.] Was it hard to ensure the correctness of the software? If there were any obstacles, what methods have been considered or practiced to improve the situation? If practiced, did it work?
\end{description}

@@

\begin{table}[H]
\centering
\hspace*{-1cm}\begin{tabular}{ll}
\hline
Threat to correctness & Num ans. \\ \hline
Complexity - data in various formats and complicated standards. & 2 \\
Complexity - different MI machines create data in different ways. & 2 \\
Complexity - additional functions beside of viewing. & 1 \\
The lack of real word image data for testing. & 1 \\
The team cannot use private data for debugging even when the data cause problems. & 1 \\
With huge datasets for testing, the tests are expensive and time-consuming. & 1 \\
It is hard to well manage releases. & 1 \\
The project has no unit tests. & 1 \\
The project has no dedicated quality assurance team. & 1 \\ \hline
\end{tabular}
\caption{\label{tab_q15_threats_correctness}Threats to correctness by the numbers of interviewees with the answers}
\end{table}

@@

\begin{table}[H]
\centering
\hspace*{-1cm}\begin{tabular}{ll}
\hline
Strategy to ensure correctness & Num ans. \\ \hline
Test-driven development, component / integration / smoke / regression tests. & 4 \\
Self tests / automated tests. & 3 \\
Two stage development process / stable release \& nightly builds. & 3 \\
CI/CD. & 1 \\
Using deidentified copies of medical images for debugging. & 1 \\
Sending beta versions to medical workers who can access the data to do the tests. & 1 \\
Collecting and maintaining a dataset of problematic images. & 1 \\ \hline
\end{tabular}
\caption{\label{tab_q15_strategies_correctness}Strategies to ensure correctness by the numbers of interviewees with the answers}
\end{table}

\subsection{Maintainability}

\begin{description}
\item[Q16.] When designing the software, did you consider the ease of future changes? For example, will it be hard to change the system’s structure, modules, or code blocks? What measures have been taken to ensure the ease of future changes and maintains?
\end{description}

@@

\begin{description}
\item[Q17.] Provide instances where users have misunderstood the software. What, if any, actions were taken to address understandability issues?
\end{description}

@@

\subsection{Usability}

\begin{description}
\item[Q17.] What, if any, actions were taken to address usability issues?
\end{description}

@@

\subsection{Reproducibility}

\begin{description}
\item[Q20.] Do you have any concern that your computational results won’t be reproducible in the future? Have you taken any steps to ensure reproducibility?
\end{description}

>>>>

Our method involves interviews with developers in the MI domain (Section \ref{sec_applying_method}). We summarize some answers from the interviews in this section. We highlight the answers that are the most informative and interesting in this section, and summarize the rest in Appendix \ref{ap_interview}.

We interviewed nine developers from eight of the 29 MI software projects. The eight projects are \textit{3D Slicer}, \textit{INVESALIUS 3}, \textit{dwv}, \textit{BioImage Suite Web}, \textit{ITK-SNAP}, \textit{MRIcroGL}, \textit{Weasis}, and \textit{OHIF}. We spent about 90 minutes for each interview and asked 20 prepared questions. We also asked following-up questions when we thought it was worth diving deeper. One participant was too busy to have an interview, so they wrote down their answers. The interviewees may have provided multiple answers to each question. Thus, when counting the number of answers, the total result is sometimes larger than nine.

@@

By asking questions \hyperlink{q9}{9}, \hyperlink{q10}{10}, and \hyperlink{q12}{12}, we tried to identify the pain points during the development process in the eight projects. The pain points include current and past obstacles. We also asked the interviewees how they would solve the problems. Questions 9, 10, and 12 are as follows:

\begin{description}
\item[Q9.] \hypertarget{q9}Currently, what are the most significant obstacles in your development process?
\item[Q10.] \hypertarget{q10}How might you change your development process to remove or reduce these obstacles?
\item[Q12.] \hypertarget{q12}In the past, is there any major obstacle to your development process that has been solved? How did you solve it?
\end{description}

@@

\begin{table}[H]
\centering
\begin{tabular}{llll}
\hline
\multirow{2}{*}{Category} & \multirow{2}{*}{Obstacle} & \multicolumn{2}{c}{Num ans.} \\ \cline{3-4} 
 &  & current & past \\ \hline
\multirow{2}{*}{Resource} & Lack of fundings & 3 &  \\
 & Lack of time to devote to the project & 2 & 1 \\ \hdashline
\multirow{3}{*}{Balance} & Hard to keep up with changes in OS and libraries & 1 &  \\
 & Hard to support multiple OS & 2 &  \\
 & Hard to support lower-end computers & 1 & 2 \\ \hdashline
Testing & Lack of access to real-world datasets for testing & 3 & 2 \\ \hdashline
\multirow{7}{*}{Others}
 & Hard to have a high level roadmap from the start & 1 &  \\
 & Not enough participants for usability tests & 1 &  \\
 & Only a few people fully understand the large codebase & 1 &  \\
 & Hard to transfer to new technologies & & 2 \\
 & Hard to understand users' needs & & 1 \\
 & Hard to maintain good documentations & & 1 \\ \hline
\end{tabular}
\caption{\label{tab_obstacles}Current and past obstacles by the numbers of interviewees with the answers}
\end{table}

The interviewees provided some potential and proven solutions for the problems in Table \ref{tab_obstacles}. We group these pain points into three major categories of obstacles: \textit{resource}, \textit{balance}, and \textit{testing}. We put the less mentioned ones into the category \textit{Others}. Sections \ref{sec_pain_points_1}, \ref{sec_pain_points_2}, and \ref{sec_pain_points_3} include further discussion about the three major groups of pain points and their solutions.

\subsection{Resource Pain Points}
\label{sec_pain_points_1}
We summarize the pain points in the \textit{resource} category as
\textbf{the lack of fundings and time.}

The potential and proven solutions are:

@@

\subsection{Balance Pain Points}
\label{sec_pain_points_2}
We summarize the pain points in the \textit{balance} category as
\textbf{the difficulty to balance between four factors: cross-platform compatibility, convenience to development \& maintenance, performance, and security.} They are also related to the choice between native application and web application.

The potential and proven solutions are:

@@

\subsection{Testing Pain Point}
\label{sec_pain_points_3}
The pain point in the \textit{testing} category is \textbf{the lack of access to real-world datasets for testing.}

The potential and proven solutions are:

@@

\noindent\textit{Proven solutions from interviewees:}
\begin{itemize}
\item asking the users to provide deidentified copies of medical images if they have problems loading the images;
\item sending the beta versions of software to medical workers who can access the data and complete the tests;
\item if (part of) the team belongs to a medical school or a hospital, using the datasets they can access;
\item if the team has access to MRI scanners, self-building sample images for testing;
\item if the team has connections with MI equipment manufacturers, asking for their help on data format problems;
\item storing all images that cause special problems, and maintaining this special dataset over time.
\end{itemize}

@@

Table \ref{tab_opinion_doc} summarizes interviewees' opinions on documentation. Interviewees from each of the eight projects thought that documentation was important to their projects, and most of them said that it could save their time to answer questions from users and developers. Most of them saw the need to improve their documentation, and only three of them thought that their documentations conveyed information clearly enough.

\begin{table}[H]
\centering
\begin{tabular}{ll}
\hline
Opinion on documentation & Num ans. \\ \hline
Documentation is vital to the project& 8 \\
Documentation of the project needs improvements & 7 \\
Referring to documentation saves time to answer questions & 6 \\
Lack of time to maintain good documentation & 4 \\
Documentation of the project conveys information clearly & 3 \\
Coding is more fun than documentation & 2 \\
Users help each other by referring to documentation & 1 \\ \hline
\end{tabular}
\caption{\label{tab_opinion_doc}Opinions on documentation by the numbers of interviewees with the answers}
\end{table}

Table \ref{tab_doc_tools} lists some of the documentation tools and methods mentioned by the interviewees.

@@

\section{Contribution Management and Project Management}
\label{sec_contribution_pm}
We tried to understand how the teams managed the contributions and their projects by asking the following questions:

\begin{description}
\item[Q5.] Do you have a defined process for accepting new contributions into your team?
\item[Q13.] What is your software development model? For example, waterfall, agile, etc.
\item[Q14.] \hypertarget{q14}What is your project management process? Do you think improving this process can tackle the current problem? Were any project management tools used?
\end{description}

Although some team may have a documented process for accepting new contributions, no one talked about it during the interview. However, most of them mentioned using GitHub and pull requests to manage contributions from the community. The interviewees generally gave very positive feedback on using GitHub. Some also said they had handled the project repository with some other tools, and eventually transferred to git and GitHub. Table \ref{tab_method_receive_contributions} shows the number of times the interviewees mentioned the methods of receiving contributions.

@@

For managing contributions, the \textit{3D Slicer} team encouraged users to develop their extensions for specific use cases, and the \textit{OHIF} team was trying to enable the use of plug-ins; the interviewee from the \textit{ITK-SNAP} team said one way of accepting new team members was through funded academic projects.

@@

Most interviewees skipped the second part of \hyperlink{q14}{Q14} ``Do you think improving this process can tackle the current problem?". In retrospect, we should not have asked a yes-or-no question, since it is not very informative. The interviewee from the \textit{OHIF} team gave a positive answer to this question. They believed that a better project management process can improve the efficiency of junior developers. They also improved the project management tools (from public Jira to public GitHub repository plus private Jira), so they could better communicate externally and internally.

\section{Discussions on Software Qualities}
\label{sec_interview_software_qualities}
Questions 15--18, and 20 are about the software qualities of \textit{correctness}, \textit{maintainability}, \textit{understandability}, \textit{usability}, and \textit{reproducibility}, respectively. We asked these questions to understand the threats to these qualities and the developers' strategies to improve them. We discuss each quality in a separate section below.

\subsection{Correctness}
\label{sec_interview_correctness}
\begin{description}
\item[Q15.] Was it hard to ensure the correctness of the software? If there were any obstacles, what methods have been considered or practiced to improve the situation? If practiced, did it work?
\end{description}

@@

\begin{table}[H]
\centering
\hspace*{-1cm}\begin{tabular}{ll}
\hline
Threat to correctness & Num ans. \\ \hline
Complexity - data in various formats and complicated standards. & 2 \\
Complexity - different MI machines create data in different ways. & 2 \\
Complexity - additional functions besides viewing. & 1 \\
The lack of real word image data for testing. & 1 \\
The team cannot use private data for debugging even when the data cause problems. & 1 \\
With huge datasets for testing, the tests are expensive and time-consuming. & 1 \\
It is hard to manage releases. & 1 \\
The project has no unit tests. & 1 \\
The project has no dedicated quality assurance team. & 1 \\ \hline
\end{tabular}
\caption{\label{tab_q15_threats_correctness}Threats to correctness by the numbers of interviewees with the answers}
\end{table}

@@

\begin{table}[H]
\centering
\hspace*{-1.5cm}\begin{tabular}{ll}
\hline
Strategy to ensure correctness & Num ans. \\ \hline
Test-driven development, component tests, integration tests, smoke tests, regression tests. & 4 \\
Self tests / automated tests. & 3 \\
Two stage development process / stable release \& nightly builds. & 3 \\
CI/CD. & 1 \\
Using deidentified copies of medical images for debugging. & 1 \\
Sending beta versions to medical workers who can access the data to do the tests. & 1 \\
Collecting and maintaining a dataset of problematic images. & 1 \\ \hline
\end{tabular}
\caption{\label{tab_q15_strategies_correctness}Strategies to ensure correctness by the numbers of interviewees with the answers}
\end{table}

\subsection{Maintainability}
\label{sec_interview_maintainability}
\begin{description}
\item[Q16.] \hypertarget{q16}When designing the software, did you consider the ease of future changes? For example, will it be hard to change the system’s structure, modules, or code blocks? What measures have been taken to ensure the ease of future changes and maintains?
\end{description}

@@

\begin{description}
\item[Q17.] \hypertarget{q17}Provide instances where users have misunderstood the software. What, if any, actions were taken to address understandability issues?
\end{description}

@@

\subsection{Usability}
\label{sec_interview_usability}
\begin{description}
\item[Q18.] \hypertarget{q18}What, if any, actions were taken to address usability issues?
\end{description}

@@

\subsection{Reproducibility}
\label{sec_interview_reproducibility}
\begin{description}
\item[Q20.] Do you have any concern that your computational results won’t be reproducible in the future? Have you taken any steps to ensure reproducibility?
\end{description}
